
---
title: "Apollo Food Forecasting"
output: R Markup
Submitted by: shikhil nangia
---

Try executing this chunk by pressing *Ctrl+Shift+Enter*. 


```{r}
dataset<-read.csv("Forecasting Food Case.csv",header = TRUE  )


#timeSeriesObj = ts(dataset$BKFST_OCCUP,start=c(2012,10,01),frequency=52)

#plot(toccupancy,ylab='Occupancy',main='Breakfast Occupancy(2012-2013)')


dateobj<- seq(as.Date("2012-10-01"), as.Date("2013-01-23"), by = "day")

ts_occupancy <- ts(dataset$BKFST_OCCUP,start=c(2012,as.numeric(format(dateobj[1], "%j"))),frequency=52)

ts_occupancy_daily <- ts(dataset$BKFST_OCCUP,start=c(2012,as.numeric(format(dateobj[1], "%j"))),frequency=365)

plot.ts(ts_occupancy,ylab='Occupancy',main='Breakfast Occupancy(2012-2013)')

```


```{r}
#install.packages('zoo')

#library(zoo)
#t1<-zoo(dataset$BKFST_OCCUP, seq(from = as.Date("2012-10-01"), to = as.Date("2013-01-23"), by = 1))
```



```{r}
cor(dataset[,c(2:9)],method = c("pearson"))

#There is high Correlation between Sambar and Chutney 

```


#Simple Moving Averages
```{r}
library(forecast)
par(mfrow = c(2,2))

lim <- c(min(dataset$BKFST_OCCUP),max(dataset$BKFST_OCCUP))

plot(dataset$BKFST_OCCUP, main="Raw Time series",ylim=lim)
plot(ma(ts_occupancy_daily,3), main="Simple moving Averages(k=3)",ylim=lim)

plot(ma(dataset$BKFST_OCCUP,7), main="Simple moving Averages(k=7)",ylim=lim)

plot(ma(dataset$BKFST_OCCUP,15), main="Simple moving Averages(k=15)",ylim=lim)

par(mfrow = c(1,1))

```

#Seasonal Decomposition by Loess Smoothing
```{r}
fit <- stl(ts_occupancy,s.window = "periodic")
plot(fit)
```




```{r}
fit_lin_trend <- tslm(ts_occupancy_daily ~ trend)
pred <- forecast(fit_lin_trend, h=2)
pred
summary(fit_lin_trend)
accuracy(fit_lin_trend)

plot(pred, xlab = "Time", ylab = "Occupancy", main = "Actual and forecasted values (Linear Trend)")

plot(ts_occupancy_daily)
lines(fit_lin_trend$fitted, col = "red", lty = 2)
points(pred$mean, col = "blue", pch = 16)

plot(fit_lin_trend$residuals)

par(mfrow = c(1,1))


```





```{r}
par(mfrow = c(2,1))
plot(ts_occupancy_daily)
lines(fit_lin_trend$fitted, col = "red", lty = 2)
points(pred$mean, col = "blue", pch = 16)

plot(fit_lin_trend$residuals)

par(mfrow = c(1,1))
```



```{r}
#Exponential Trend
fit_exp_trend <- tslm(ts_occupancy_daily ~ trend, lambda = 0)
#lambda=0 means conversion of y (sales) into log 
pred <- forecast(fit_exp_trend, h=4)
pred

accuracy(fit_exp_trend)

plot(pred, xlab = "Time", ylab = "Occupancey", main = "Actual and forecasted values")

par(mfrow = c(2,1))
plot(ts_occupancy_daily)
lines(fit_exp_trend$fitted, col = "red", lty = 2)
points(pred$mean, col = "blue", pch = 16)

plot(fit_exp_trend$residuals)

par(mfrow = c(1,1))
```

##Checking for Quadratic Trend
```{r}
fit_quad_trend <- tslm(ts_occupancy_daily ~ trend + I(trend^2))

#Forecasting for 2 periods
pred <- forecast(fit_quad_trend, h=2)
pred

accuracy(fit_quad_trend)

plot(pred, xlab = "Time", ylab = "Occupancy", main = "Actual and forecasted values")

plot(ts_occupancy_daily)
lines(fit_quad_trend$fitted, col = "red", lty = 2)
points(pred$mean, col = "blue", pch = 16)

```

##Model with Seasonality
```{r}
fit_season <- tslm(ts_occupancy_daily ~ season)

pred <- forecast(fit_season, h=2)
pred

accuracy(fit_season)

plot(pred, xlab = "Time", ylab = " Occupancy", main = "Actual and forecasted values")

```

#Trend and Seasonality
```{r}

fit_season_trend <- tslm(ts_occupancy_daily ~ trend + I(trend^2) + season)

pred <- forecast(fit_season_trend, h=2)
pred

accuracy(fit_season_trend)

plot(pred, xlab = "Time", ylab = "Occupancy", main = "Actual and forecasted values")
```


#ARIMA
```{r}

ndiffs(ts_occupancy_daily)

##Outputs 0 , however we will check with ADF TEst


dOccupancy <- diff(ts_occupancy_daily)

plot(dOccupancy)

# Apply ADF test to check for stationarity.

adf.test(dOccupancy)

# ADF test proves that time series is stationary, p value : 0.01

```


```{r}
par(mfrow = c(2,1))

Acf(dOccupancy)
Pacf(dOccupancy)

par(mfrow = c(1,1))
```
#Fitting the model
```{r}
fit <- arima(dOccupancy, order = c(0,1,1))
fit
accuracy(fit)

```


#### Automated ARIMA Forecasting 
```{r}

fit1 <- auto.arima(ts_occupancy_daily)
fit1

forecast(fit1, 3)

accuracy(fit1)

plot(forecast(fit1, 3))
```



#Exponential smoothing assuming only level in series
```{r}
ANN <- ets(ts_occupancy_daily, model = "ANN")
ANN

pred_ANN <- forecast(ANN, 2)
pred_ANN

plot(pred_ANN, xlab = "Year", ylab = "Occupancy", main = "Breakfast Occupied")

accuracy(ANN)
```

#Exponential smoothing assuming level and trend in the series
```{r}
AAN <- ets(ts_occupancy_daily, model = "AAN")
AAN

pred_AAN <- forecast(AAN, 2)
pred_AAN

plot(pred_AAN, xlab = "Year", ylab = "Occupancy", main = "Brkfast Occupancy")

accuracy(AAN)
```

#Exponential smoothing assuming level, trend and seasonality
```{r}
AAA <- ets(ts_occupancy_daily, model = "AAA")
AAA

pred_AAA <- forecast(AAA, 2)
pred_AAA
plot(pred_AAA, xlab = "Time", ylab = "Occupancy", main = "Brkfast Occupancy")

accuracy(AAA)

```

#Automatic exponential forecasting with ets ()
```{r}
Automatic <- ets(ts_occupancy_daily)
Automatic

plot(forecast(Automatic), main = "Bkfst Occupancy", xlab = "Time", ylab = "Occupancy", flty = 2)

accuracy(Automatic)


```
# Training Test Partition
```{r}
time = time(ts_occupancy_daily)
stepsAhead <- 14   #VALIDATION FOR 14 days
nTrain <- length(ts_occupancy_daily) - stepsAhead
train.ts <- window(ts_occupancy_daily, start = time[1], end = time[nTrain])
valid.ts <- window(ts_occupancy_daily, start = time[nTrain + 1], end = time[nTrain + stepsAhead])
```
#Forecast using Linear Trend
```{r}
train.lm <- tslm(train.ts ~ trend)
train.lm.pred <- forecast(train.lm, h = stepsAhead)
accuracy(train.lm.pred, valid.ts)
```


